Sequential rank agreement methods for comparison of ranked lists

### 2017-04-07 latex/superranker-manuscript.tex
forkortelse af simu sektion

Lasso shows similar agreement for ranking the risk predictions than
PLS-DA (unless with the AUC criterion), and poorer agreement for ranking
predictors. This reason for the latter is the high auto-correlation
between the intensities in the mass spectra which leads to
collinearity issues in the regression models. It is well-known that
variable selection with the Lasso penalty does not preform very well
when the predictors are highly correlated. The collinearity does,
however, not affect the agreement of the risk predictions, as seen on
the left panel of Figure \ref{fig:app1}, since it is not so important
which specific variable that gets selected from a group of highly correlated
predictors when the purpose is risk predictions.

@@

>>>>

Lasso shows similar agreement for ranking the risk predictions than
PLS-DA (except for the AUC criterion), and poorer agreement for
ranking predictors. This reason for the latter is the high
auto-correlation between the intensities in the mass spectra which
leads to collinearity issues in the regression models. It is
well-known that variable selection with the Lasso penalty does not
preform very well when the predictors are highly correlated. The
collinearity does, however, not affect the agreement of the risk
predictions, as seen on the left panel of Figure \ref{fig:app1}, since
it is not so important which specific variable that gets selected from
a group of highly correlated predictors when the purpose is risk
predictions.

@@

\section{Simulation study --- comparison of list agreements}

We present results of a simulation study where we investigated the
small sample properties of our proposed method (sra) and compared it
with the topK method proposed by \citet{hall:schi:2012} with respect
to list depth agreement. In each simulation run, we generated a
dataset with 1000 predictors and 400 observations. The predictors were
drawn independently from a standard Gaussian distribution. The outcome
was drawn from a Gaussian distribution with variance 1 and mean that
depended additively on the first 15 predictors, i.e.,

$$y_i = \sum_{j=1}^{15}x_{ij} + \epsilon_i,$$

where $y_i$ is the $i$th response, and $x_{ij}$ is the $j$th predictor
for the $i$th measurement, and $\epsilon_i\sim N(0, 1)$. In each
simulation run we obtained \(L\) ranked lists of the 1000 predictors.
This was done by first drawing \(L\) bootstrap samples (with
replacement) and then in each bootstrap sample ranking the 1000
predictors according to their marginal $t$ test statistics. Thus, the
simulation results compare lists that are obtained with the same
technique on bootstrap versions of the same dataset. We report results
from two sets of simulations (each based on 1000 simulated datasets):
\begin{enumerate}
\item[] Scenario I: \(L=8\) and $q\in\{3,4,5,6,7,8,9,10\}$.
\item[] Scenario II: \(q=5\) and $L\in\{3,5,10,50\}$.
\end{enumerate}
In both scenarios, we summarized the distribution of the estimated
depth of agreement and the average number of different predictors
found at a given depth.
 % Asymptotic results are available for both methods as the number of
 % lists increases, so this will investigate how much the results might
 % change as $L$ increases. For a fixed constant threshold of $q_L=5$ we
 % increase the number of available lists to investigate the sensitivity
 % of the results to $L$.
  % Throughout the simulations we
  % assume a constant, prespecified threshold function, $q_L$, and try
  % different values for the constant. The approach by
  % \citet{hall:schi:2012} uses a window size parameter, $K$, to
  % represent the maximum difference in rankings before they consider an
  % item in two lists to be in agreement.  Because the standard
  % deviation --- which is used in the sequential rank agreement
  % definition of $\hat{A}_L$--- squares its differences, it tends to
  % give more weight to larger rank differences and less weight to
  % smaller differences compared to the mean absolute difference.  To
  % ease comparison between the sequential rank agreement and the
  % \citet{hall:schi:2012} approach we would like to use the same
  % threshold for both methods, and in order to do this we use the
  % median absolute deviance throughout the simulations in lieu of the
  % pooled standard deviation.
 % For all parameter combinations we ran 1000 simulations and report the
 % distribution of the estimated depth of agreement and the average
 % number of different predictors found at a given depth.
The results from Scenario I are shown in the left panel of
Figure~\ref{fig:example99}. The double histograms show the
distribution of the estimated depth of agreement. We see that sra
results in a substantially lower depth of agreement than the topK
method. Also the average numbers of unique predictors (bold numbers in
the double histograms) are markedly smaller for sra. Even larger
differences are found if we used the Euclidean distance instead of the
median absolute distance for the sequential rank agreement (results
not shown). As expected the depth of agreement increases when the
threshold for agreement/window increases.
\begin{figure}[htbp]
   \begin{center}
 \includegraphics[width=.49\textwidth]{Ksim.pdf}
 \includegraphics[width=.49\textwidth]{Lsim}
 \end{center}
 \caption{Left panel: Simulation study showing distribution of
   estimated rank agreements for sra and topK for varying thresholds
   and fixed number of lists $L=8$. The bold numbers are the average
   number of unique predictors in the set found to be in
   agreement. Right panel: Simulation results for varying number of
   lists, and with fixed threshold of $q=5$.}
 \label{fig:example99}
 \end{figure}
 The right panel of Figure~\ref{fig:example99} shows the results from
 Scenario II. The number of lists has little impact but at all depths
 sra is more conservative than topK. As a consequence we also see that
 sra systematically selects fewer predictors in the parts of
 agreement.
